---
title: "Covid19 vaccine study Fall2021"
author: "Abel Jacobo Perez"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 5

---





```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalData <-read.csv(file="fall21.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
guy = filter(FinalData, sex =="guy")
gal = filter(FinalData, sex =="gal")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```
# Introduction

A new  preventative vaccine (covid21) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Part of the information you need to compute is the so-called effectiveness of the drug for each subgroup of people. Include a 95% confidence interval of the effectiveness for each different subgroup. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

$H_0$:There is no difference in  covid-19 vaccine effectiveness amongst difference subgroups(males,females,drug users, LGBTQ)

$H_A$:There is a difference in  covid-19vaccine effectiveness amongst different subgoups(males,females,Drug users,LGBTQ)


# Males

### Barplot

```{r}
dd2 <- guy %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```






This plot represents the population of guys(males) people and the relationship to the contraction of covid-19 by either talking the vaccine or the placebo. 


The plot above shows that a shocking 58% of the people that took the vaccine did in fact get covid-19, while only approximately 42% of the people taking the placebo contracted covid-19.  A substantially higher percentage of people who took the vaccine contracted covid-19 versus those who were only given the placebo. The vaccine appears to be ineffective in protection against covid-19. The odds of getting covid-19 while taking the vaccine is 3:2. The men that took the drug were one and half times more likely to get covid-19 than those taking the placebo.


There are two plots. In first plot,the x-axis represents infected, whether the participants were infected with covid-19 or they were not. The y-axis represents the count, or number of people that got covid-19 or not.
In the second plot, the one above, The The left two bars of the graph are all the people that got covid-19. The x-axis represents whether a person got covid-19 or not. The y-axis, however, instead of representing the number of guys who got covid-19, now represents the percentage of guys that got covid-19 after taking the the placebo or taking the Vaccine. 


The codes that  create this plot start with the code block beginning with dd2. dd2 can be seen in the environment pane and if inspected contains 4 variables.These variables include infected, treatment, count, and prcent(stands for percent). The second group of codes is called the basicC. BasicC can also be seen in the environment pane and appears to contain a list of 9 variables. This information creates the first plot. The basicCC is the last bunch of codes and is also in the files pane. It also contains 9 variables and helps to create the percentage plot, the second plot.


### Numerical results

  


```{r}
table1 <- xtabs(~infected + treatment, data=guy)
rowPerc(table1)
colPerc(table1)
```
These numerical results show specific percentages of guys(men) that got covid-19 after taking the placebo and taking the vaccine. 
Among the guys, 58% got covid-19 after taking the vaccine, while the approximately 42% of guys that were given the placebo contracted covid-19. The vaccine is clearly not effective in preventing covid-19 in guys(males) and it is not recommended. Those who took the placebo were less likely to contact covid-19
The odds of getting covid-19 while taking the vaccine is 3:2. The men that took the drug were one and half times more likely to get covid-19 than those taking the placebo.

### Inferential results

```{r}
chisq.test(table1)
chisqtestGC(table1)
guyfish <- fisher.test(table1)
guyfish
```




The information to play close attention to among all of these results is the output of the Fisher's Exact test. The p-value displayed above is extremely small at 6e-04. This is an awfully small number. seeing this number means that these results are not likely to occur by chance.



The null hypothesis predicted that the deaths related to covid-19 are independent of Subgroup. Since it is clear that Subgroups do play a role in the covid-19 infection rate of individuals, the alternative hypothesis, that there is a correlation between covid-19 vaccine effectiveness and subgroups, is the favorable option.

The odd ratio above illustrates a ratio of 0.68. the guys(men) who received the vaccine were 1.5 times more likely to get covid-19 than those who received the placebo. 


```{r}
effect <- guyfish$p.value/(guyfish$p.value +1)
effectivenessguy = guyfish$estimate/(guyfish$estimate +1)*100
print(effectivenessguy)
```

```{r}
low=(guyfish$conf.int[1]/(guyfish$conf.int[1] +1))*100
hi=(guyfish$conf.int[2]/(guyfish$conf.int[2] +1))*100
options(digits=2)
```

Above is the effectiveness for the covid-19 vaccine for guys(males). If the effectiveness is above 50% that means that the vaccine is effective. Based on the interval seen above, we can conclude that the vaccine is ineffective in protecting guys from getting covid-19. 

The odds ratio is approximately how effective the vaccine is. The odds ratio for the guys can be seen above as 40. The effectiveness confidence interval for the guys is 35.09 to 45.94.

# Females



```{r}
dd2 <- gal %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```






This plot represents the population of girls(females)and the relationship to the contraction of covid-19 by either talking the vaccine or the placebo. 


The plot above shows that only 33% of the girls that took the vaccine did in fact get covid-19, while 67% of the people taking the placebo contracted covid-19.  A substantially higher percentage of people who took tha vaccine contracted covid-19 versus those who were only given the placebo. The vaccine appears to be effective in protection against covid-19 for girls(females). The odds of getting covid-19 while taking the vaccine is 1:3 The woman that took the drug were half as likely to get covid-19 than those taking the placebo.


There are two plots. In first plot,the x-axis represents infected, whether the participants were infected with covid-19 or they were not. The y-axis represents the count, or number of people that got covid-19 or not.
In the second plot, the one above, The The left two bars of the graph are all the woman that got covid-19. The x-axis represents whether a person got covid-19 or not. The y-axis, however, instead of representing the number of woman who got covid-19, now represents the percentage of woman that got covid-19 after taking the the placebo or taking the Vaccine. 


The codes that  create this plot start with the code block beginning with dd2. dd2 can be seen in the environment pane and if inspected contains 4 variables.These variables include infected, treatment, count, and prcent(stands for percent). The second group of codes is called the basicC. BasicC can also be seen in the environment pane and appears to contain a list of 9 variables. This information creates the first plot. The basicCC is the last bunch of codes and is also in the files pane. It also contains 9 variables and helps to create the percentage plot, the second plot.


### Numerical results

  


```{r}
table1 <- xtabs(~infected + treatment, data=gal)
rowPerc(table1)
colPerc(table1)
```
These numerical results show specific percentages of girls(females) that got covid-19 after taking the placebo and taking the vaccine. 
Among the gals, 33% got covid-19 after taking the vaccine, while the approximately 67% of girls that were given the placebo contracted covid-19. The vaccine is clearly effective in preventing covid-19 in girls(females) and it is not recommended. Those who took the placebo were much less likely to contact covid-19
The odds of getting covid-19  for woman while taking the vaccine is 1:2. The women that took the drug were half as likely to get covid-19 (33%) than those taking the placebo(67%).

### Inferential results

```{r}
chisq.test(table1)
chisqtestGC(table1)
galfish <- fisher.test(table1)
galfish
```




The information to play close attention to among all of these results is the output of the Fisher's Exact test. The p-value displayed above is extremely small at 2e-16. This is an awfully small number. This number means that these results are not likely to occur by chance.



The null hypothesis predicted that the infection rate related to covid-19 vaccination effectivness are independent of subgroup. Since it is clear that subgroup does play a role in the Covid infection rate in the vaccinated individuals, the alternative hypothesis, that there is a correlation between covid-19 vaccination effectiveness and subgroups, is the favorable option.

The odd ratio above illustrates a ratio of 2.00. the women(females) who received the vaccine were half as likely to get covid-19 than those who received the placebo. 


```{r}
effect <- galfish$p.value/(galfish$p.value +1)
effectivenessgal = galfish$estimate/(galfish$estimate +1)*100
print(effectivenessgal)
```

```{r}
low=(galfish$conf.int[1]/(galfish$conf.int[1] +1))*100
hi=(galfish$conf.int[2]/(galfish$conf.int[2] +1))*100
options(digits=2)
```

Above is the effectiveness for the covid-19 vaccine for woman(females). If the effectiveness is above 50% that means that the vaccine is effective. Based on the interval seen above, we cn conclude that the vaccine is effective in protecting woman(females) from getting covid-19. 

The odds ratio is approximately how effective the vaccine is. The odds ratio for the woman(females) can be seen above as 67. The effectiveness confidence interval for the females is 55.09 to 75.05

# Drug Users


```{r}
dd2 <- Druguser %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```






This plot represents the population of drug users and the relationship to the contraction of covid-19 by either talking the vaccine or the placebo. 


The plot above shows that only 33% of the girls that took the vaccine did in fact get covid-19, while 67% of the drug users taking the placebo contracted covid-19.  A substantially lower percentage of people who took the vaccine contracted covid-19 versus those who were only given the placebo. The vaccine appears to be effective in protection against covid-19 for Drug users. The odds of getting covid-19 while taking the vaccine is 1:3 The drug users that took the drug were half as likely to get covid-19 than those taking the placebo.


There are two plots. In first plot,the x-axis represents infected, whether the participants were infected with covid-19 or they were not. The y-axis represents the count, or number of drug users that got covid-19 or not.
In the second plot, the one above, The The left two bars of the graph are all the drug users that got covid-19. The x-axis represents whether the person got covid-19 or not. The y-axis, however, instead of representing the number of drug users who got covid-19, now represents the percentage of drug users that got covid-19 after taking the the placebo or taking the Vaccine. 


The codes that  create this plot start with the code block beginning with dd2. dd2 can be seen in the environment pane and if inspected contains 4 variables.These variables include infected, treatment, count, and prcent(stands for percent). The second group of codes is called the basicC. BasicC can also be seen in the environment pane and appears to contain a list of 9 variables. This information creates the first plot. The basicCC is the last bunch of codes and is also in the files pane. It also contains 9 variables and helps to create the percentage plot, the second plot.


### Numerical results

  


```{r}
table1 <- xtabs(~infected + treatment, data=Druguser)
rowPerc(table1)
colPerc(table1)
```
These numerical results show specific percentages of drug users that got covid-19 after taking the placebo and taking the vaccine. 
Among the Drug users, 33% got covid-19 after taking the vaccine, while the approximately 67% of drug users that were given the placebo contracted covid-19. The vaccine is clearly effective in preventing covid-19 in drug users and it is recommended. Those who took the placebo were much more likely to contact covid-19
The odds of getting covid-19 while taking the vaccine is 1:2. The drug users that took the drug were half as likely to get covid-19 (33%) than those taking the placebo(67%).

### Inferential results

```{r}
chisq.test(table1)
chisqtestGC(table1)
Druguserfish <- fisher.test(table1)
Druguserfish
```




The information to play close attention to among all of these results is the output of the Fisher's Exact test. The p-value displayed above is extremely small at .003 This is an awfully small number. This number means that these results are not likely to occur by chance.



The null hypothesis predicted that the covid-19 vaccine effectiveness is independent of subgroup. Since it is clear that subgroup does play a role in the rate of infection in individuals who were vaccinated, the alternative hypothesis, that there is a correlation between covid-19 infection rate in those vaccinated and subgroups, is the favorable option.

The odd ratio above illustrates a ratio of 1.8. the Drug users who received the vaccine were half as likely to get covid-19 than those who received the placebo. 


```{r}
effect <- Druguserfish$p.value/(galfish$p.value +1)
effectivenessDruguser = Druguserfish$estimate/(Druguserfish$estimate +1)*100
print(effectivenessDruguser)
```

```{r}
low=(Druguserfish$conf.int[1]/(Druguserfish$conf.int[1] +1))*100
hi=(Druguserfish$conf.int[2]/(Druguserfish$conf.int[2] +1))*100
options(digits=2)
```

Above is the effectiveness for the covid-19 vaccine for drug users. If the effectiveness is above 50% that means that the vaccine is effective. Based on the interval seen above, we can conclude that the vaccine is effective in protecting drug users from getting covid-19. 

The odds ratio is approximately how effective the vaccine is. The odds ratio for the drug users can be seen above as 65. The effectiveness confidence interval for the Drug users is 55.09 to 75.05

# LGBTQ

# Conclusions and recommendations

